Cover Times of Random Walks

### 2019-04-08 paper.tex
typo fixes

Let $T_{j}$ be the number of steps until the first visit to $j$.
Formally,
\begin{align}
T_{j} &= \textrm{min} \left\{t : X_t = j \right\}. \nonumber
\end{align}
Let the hitting time $\hit$ be the maximum expected time between two vertices on $G$.
Formally, 
\begin{align}
\hit &= \max_{i,j \in V(G)} \E_i[T_j]. \nonumber
\end{align}
Note that $\E_i[T_j] = \E[T_j | X_0 = i]$.
Let the cover time $\cov$ be the until all vertices have been
visited by the random walk $(X_t)$.
We typically refer to the expected cover time $\cov$.
Formally,
\begin{align}
\cov &= \E [ \max_{j \in V(G)} T_j ]. \nonumber
\end{align}

@@

\cref{fig:example} shows our small graph and the first step analysis
for the random walk $(X_t)$ started from the top left vertex.
The black vertex indicates the current vertex of the random walk
and the empty vertices indicate the visited nodes.

For a graph with $n$ vertices, there are $n$ possible choices for the current
vertex and $2^{n-1}$ possible combinations of visited and unvisited vertices
for each one.
Of course, some states are inaccessible.
But even on \cref{fig:example} where we exploit symmetry,
there are almost a prohibitive number of cases.
It follows from our combinatorial analysis that the number of states
grows exponentially with the number of vertices.

@@

	\caption{First step analysis applied to an asymmetric graph with $n=4$.}\label{fig:example}

@@

\subsection{Complete Graph}
We call the graph with $n$ vertices and ${n \choose 2}$ edges the complete graph.

@@

The random variable $X_i$ takes value $k$ when $k-1$ steps lead us to
already visited vertices and the $k^{th}$ step is to previously unvisited vertex.
It follows that $X_i$ is geometric with distribution
\begin{align}
\pr(X_i = k) = (1-p_i)^{k-1}p_i, \nonumber
\end{align}
where $p_i$ is the probability of moving to a previously unvisited vertex.
Then $\E[X_i] = \frac{1}{p_i}$. We now find $p_i$:
the probability that we go to a previously unvisited vertex is $\frac{n-i}{n-1}$.
This is because there are $n-i$ unvisited vertices and a total of $n-1$ adjacent vertices.
It follows that $\E[X_i] = \frac{n-1}{n-i}$.
Then
\begin{align}
\cov&=   \sum_{i=1}^n \frac{n-1}{n-i} \nonumber \\
&=  (n-1) \left(\frac{1}{n-1}+\frac{1}{n-2}+...+1\right). \nonumber
\end{align}

@@

To solve the telescoping equations, observe that the walk is symmetric:
we could flip the walk over the vertical axis without changing $e_i$.
This makes sense because our intuition tells us that the only 
identifying feature of $i$ is its respective distances to the endpoints.
The logical conclusion is that $e_i=e_{N-i}$.
In particular, we have already seen that $e_0=e_N=0$ and it
naturally follows that $e_1=e_{N_1}$.
We use these observations to find $e_1$.

@@

In a slight abuse of notation, let $T_k$ be the 
number of steps until we reach $k$ distinct vertices.
Formally, $T_k = \min{t: b-a+1 = k}$.
Then by telescoping,
\begin{align}
\cov &= \E[T_n] = \E[(T_2 - T_1) + (T_3 - T_2) + \dots + 
(T_k - T_{k-1}) + \dots + (T_n - T_{n-1})] \nonumber
\end{align}

>>>>

Let $T_{j}$ be the number of steps until the first visit to $j$.
Formally,
\begin{align}
T_{j} &= \textrm{min} \left\{t : X_t = j \right\}. \nonumber
\end{align}
Let the hitting time $\hit$ be the maximum expected time between two vertices on $G$.
Formally, 
\begin{align}
\hit &= \max_{i,j \in V(G)} \E_i[T_j]. \nonumber
\end{align}
Note that $\E_i[T_j] = \E[T_j | X_0 = i]$.
Let the cover time be the time until all vertices have been
visited by the random walk $(X_t)$.
We typically refer to the expected cover time $\cov$.
Formally,
\begin{align}
\cov &= \E [ \max_{j \in V(G)} T_j ]. \nonumber
\end{align}

@@

\cref{fig:example} shows our small graph and the first step analysis
for the random walk $(X_t)$ started from the top left vertex.
The black circle indicates the current vertex of the random walk
and the empty circles indicate the visited vertices.

For a graph with $n$ vertices, there are $n$ possible choices for the current
vertex and $2^{n-1}$ possible combinations of visited and unvisited vertices
for each one.
Of course, some states are inaccessible.
But even on \cref{fig:example} where we exploit symmetry,
there are almost a prohibitive number of cases.
It is easy to see that the number of states
grows exponentially with the number of vertices.

@@

	\caption{First step analysis applied to a small, asymmetric graph with $n=4$.}\label{fig:example}

@@

\subsection{Complete Graph}
We call the graph with an edge between each pair of vertices the
complete graph.
Note that a complete graph with $n$ vertices has ${n \choose 2}$ edges.

@@

The random variable $X_i$ takes value $k$ when $k-1$ steps lead us to
already visited vertices and the $k^{th}$ step is to a previously unvisited vertex.
It follows that $X_i$ is geometric with distribution
\begin{align}
\pr(X_i = k) = (1-p_i)^{k-1}p_i, \nonumber
\end{align}
where $p_i$ is the probability of moving to a previously unvisited vertex.
Then $\E[X_i] = \frac{1}{p_i}$. We now find the probability $p_i$ that 
we go to a previously unvisited vertex is $\frac{n-i}{n-1}$.
This is because there are $n-i$ unvisited vertices and a total of $n-1$ adjacent vertices.
It follows that $\E[X_i] = \frac{n-1}{n-i}$.
Then
\begin{align}
\cov&=   \sum_{i=1}^n \frac{n-1}{n-i} \nonumber \\
&=  (n-1) \left(\frac{1}{n-1}+\frac{1}{n-2}+...+1\right). \nonumber
\end{align}

@@

To solve the telescoping equations, observe that the walk is symmetric:
we could flip the walk over the vertical axis without changing $e_i$.
This makes sense because our intuition tells us that the only 
identifying feature of $i$ is its respective distances to the endpoints.
The logical conclusion is that $e_i=e_{N-i}$.
In particular, we have already seen that $e_0=e_N=0$ and it
naturally follows that $e_1=e_{N_1}$.
We use these observations to find $e_1$:

@@

In a slight abuse of notation, let $T_k$ be the 
number of steps until we reach $k$ distinct vertices.
Formally, $T_k = \min\{t: b-a+1 = k\}$.
Then by telescoping,
\begin{align}
\cov &= \E[T_n] = \E[(T_2 - T_1) + (T_3 - T_2) + \dots + 
(T_k - T_{k-1}) + \dots + (T_n - T_{n-1})] \nonumber
\end{align}
